PANCHANGE_data_preprocessing

### 2020-08-11 PANCHANGE_data_preprocessing2020-07-15.Rmd
key worker chunk

>>>>

# GLAD employment
## Read in data
```{r glad dem cols}
if(GLAD == TRUE){
glad.employment.raw <- readRDS(file = paste0(data.raw_path, "/glad/employment_coping_glad.rds"))
dim(glad.employment.raw)
colnames(glad.employment.raw)
}
```

# Select variables
```{r glad employment cleaning}
if(GLAD == TRUE){
exclude_cols_employment <- c("ID",
                  "Sample")

glad.employment.raw.id <- glad.employment.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(
    ID = as.numeric(ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers
         ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(glad.employment.raw.id)
# Inspect colnames
colnames(glad.employment.raw.id)
#Differences
dim(glad.employment.raw)[1]-dim(glad.employment.raw.id)[1]
}
```

# Highest education
Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
Assumes University degree > A level > NVQ > GCSE/O Level/CSE
```{r Highest education}
if(GLAD == TRUE | EDGI == TRUE) {
#create numeric version of the highest education variable
glad.employment.raw.id <- glad.employment.raw.id %>%
  mutate(
    highest_education_numeric =
      case_when(
        employment.phd_numeric == "1" ~ 4,
        employment.masters_degree_or_equivalent_numeric == "1" ~ 4,
        employment.postgraduate_degree_or_equivalent_numeric == "1" ~ 4,
        employment.bachelors_degree_or_equivalent_numeric == "1" ~ 4,
        employment.a_levelsas_levels_or_equivalent_numeric == "1" ~ 3,
        employment.nvq_or_hnd_or_hnc_or_equivalent_numeric == "1" ~ 2,
        employment.o_levelsgcses_or_equivalent_numeric == "1" ~ 1,
        employment.cses_or_equivalent_numeric == "1" ~ 1
      )
    )

#recode the numeric version into a factor
glad.employment.raw.id <- glad.employment.raw.id %>%
  mutate(
    highest_education =
      recode_factor(
        highest_education_numeric,
        `1` = "GCSE/CSE",
        `2` = "NVQ",
        `3` = "A-levels",
        `4` = "University")
    )

glad.employment.raw.id %>%
  freq(highest_education)
}
```

Check highest education coding
```{r Highest education check}
if(GLAD == TRUE | EDGI == TRUE) {
glad.employment.raw.id %>%
  select(highest_education,
         employment.phd_numeric,
         employment.masters_degree_or_equivalent_numeric,
         employment.postgraduate_degree_or_equivalent_numeric,
         employment.bachelors_degree_or_equivalent_numeric,
         employment.a_levelsas_levels_or_equivalent_numeric,
         employment.nvq_or_hnd_or_hnc_or_equivalent_numeric,
         employment.o_levelsgcses_or_equivalent_numeric,
         employment.cses_or_equivalent_numeric
  )
}
```

# Key worker status
```{r Key worker}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE | RAMP == TRUE) {
#create numeric version of the variable
glad.employment.raw.id <- glad.employment.raw.id %>%
  mutate(
    key_worker_numeric =
      case_when(
        employment.government_work_key_workers_numeric == "0" ~ 0,
        employment.government_work_key_workers_numeric > 0 ~ 1
      )
    )

#recode the numeric version into a factor
glad.employment.raw.id <- glad.employment.raw.id %>%
  mutate(
    key_worker =
      recode_factor(
        key_worker_numeric,
        `0` = "No key worker",
        `1` = "Key worker"
        )
    )

glad.employment.raw.id %>%
  freq(key_worker)
}
```
